Block bootstrap consistency\\under weak assumptions

### 2017-02-05 Paper.tex
[#35/gfn] Address referee's snarky comment on functional assumptions
[#40/check] Tweak LaTeX code after lacheck

The assumption on the dispersion of the individual means,
$(\mu_{nt} - \bar\mu_n)^2$, is the same as \citepos{GoW:02} and
\citepos{GoJ:03} under their additional requirement that the
expected block length be $o(\sqrt{n})$. When
the blocks are larger than $o(\sqrt{n})$, a case that \citet{GoW:02}
and \citet{GoJ:03} do not consider, the dispersion must be smaller
than they require.
To see these implications of this assumption, it is helpful to
consider a simple example: a stochastic process with a single break of
size $\Delta_n$ at time $T_n$:
\begin{equation*}
\mu_{nt} =
\begin{cases}
  \mu & \text{if } t \leq T_n \\
  \mu + \Delta_n & \text{otherwise}.
\end{cases}
\end{equation*}
A small amount of algebra gives the relationship,
\begin{equation*}
  \sum_{t=1}^n (\mu_{nt} - \bar \mu_n)^2
  = T_n (n - T_n) \Delta_n^2 / n \equiv D_n;
\end{equation*}
we have introduced $D_n$ to minimize notation in the subsequent
discussion.  For the Moving and Circular Block Bootstraps,
Theorem~\ref{T1}'s Assumption 2 holds if $D_n = o(n / M_n^2)$, and
the same relationship must hold for the expected block length for the
Stationary Bootstrap. If $\Delta_n = \Delta_0$ and $T_n = \pi n$ for
some constants $\Delta_0$ and $\pi$ and for all $n$, then
$D_n \approx \Delta_0^2 \, \pi (1 - \pi) \, n$ and Assumption~2 fails to
hold. These values of $\Delta_n$ and $T_n$ describe a scenario where
there is a break large enough that $\pi$ could be precisely
estimated \citep[as in][]{Bai:94}, but the break is ignored and the
bootstrap is applied to the entire dataset.

To see when Theorem~\ref{T1}'s Assumption~2 is satisfied, consider two
other parametrizations of this DGP\@. First, let $\Delta_n = \Delta_0$
and $T_n = T_0$ for some finite $T_0$. This mimics a setting where a
researcher intends to apply the bootstrap to a post-break sample, but
mistakenly includes a small number of pre-break observations as
well. In this case, Assumption 2 holds as long as
$M_n = o(\sqrt{n})$.  Similarly, suppose
$\Delta_n = \Delta / \sqrt{n}$ and $T_n = \pi n$, which mimics a
setting with a moderate break with unknown timing.  Then
$D_n \approx \Delta \pi (1 - \pi)$ and Assumption 2 again holds as long as
$M_n = o(\sqrt{n})$. Consequently, using block sizes of length
$o(\sqrt{n})$ ensures that Theorem~\ref{T1}'s assumptions on the dispersion
of the $\mu_{nt}$ hold when there is a moderate amount of
heterogeneity---either because the sample is chosen to lie after a
large estimated break or because any breaks are smaller in
magnitude and, consequently, potentially difficult to detect. When the
degree of heterogeneity is very small or nonexistent, so
$\sum_t (\mu_{nt} - \bar \mu_n)^2$ grows even slower, then larger
block lengths can also satisfy this condition.

@@

Our insight is that the expectation of squared blocks of the bootstrap
process can be expressed as a sequence of \emph{contiguous} blocks of
the original process, so the arguments that establish convergence of
the original squared blocks can be applied with only minor changes to
the bootstrapped blocks. Consider the Moving Blocks Bootstrap,%
\footnote{To make this presentation as simple as possible, assume for
  now that $n = M_n J_n$ exactly.} %
for example, and let
\begin{equation*}
  Z_{nj}^* = \tfrac{1}{\sqrt{n}} \sum_{t=(j-1) M_n + 1}^{j M_n} (X_{nt}^* - \bar X_n).
\end{equation*}
Then, conditional on the data, the $Z_{nj}^{*2}$ are independent and can
be expected to obey an \lln, so %
$\sum_j (Z_{nj}^{*2} - \E^* Z_{nj}^{*2}) \to^p 0$ %
and the CLT for the bootstrapped array requires $\sum_j \E^*
Z_{nj}^{*2}$ to converge to a positive and finite limit.  But, since
$\E^*$ only averages over the starting point of each block, we have
\begin{align*}
  \E^* Z_{nj}^{*2} &=
  \tfrac{1}{n} \sum_{\tau = 0}^{n-M_n}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= \tau + 1}^{\tau + M_n} (X_{nt} - \bar X_n) \Big)^2 \\
  &= \tfrac{1}{n} \sum_{\tau_0 = 0}^{M_n-1}
  \sum_{j=1}^{J_n}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= (j - 1) M_n + \tau_0 + 1}^{j M_n + \tau_0} (X_{nt} - \bar X_n) \Big)^2
\end{align*}
after grouping blocks separated by $M_n$ periods. For each $\tau_0$,
the summation can be expected to converge in probability through the
same arguments that were used to establish the \clt\ for the original
array.%
\footnote{Some of the details of the argument will typically need to
  change because the original \clt\ only requires convergence for
  $\tau_0 = 0$, but these details are often incidental to the original
  argument.}%
\footnote{Lemmas~\ref{L4} and~\ref{L6} are particularly
  strong demonstrations of this argument.} %
A similar representation is available for the Circular and Stationary
bootstraps.

@@

First, the relevant probability infrastructure.  Let
$(S, \Omega, \pr)$ be a probability space and define the sequence of
sub-sigma-fields $\Omega_n \subset \Omega$. Assume that each
$X_{nt}$, $V_{nt}$, $M_{nj}$, and $u_{nj}$ is $\Omega_n$-measurable,
with $u_{nj}$ the $\operatorname{uniform}(1,\dots,n)$ random variables
that designate the start period of each bootstrap ``block.'' Also
define the \sigmafield\ generated by the stationary bootstrap's block
lengths alone,
\begin{equation}
  \Ms = \sigma(J_{n}, M_{n1},\dots,M_{nJ_n})
\end{equation}
and the conditional probability $\Pm(\cdot) = \pr[\cdot \mid \Ms]$.
(And define $\Pm^*(\cdot) = \pr^*[\cdot \mid \Ms]$, $\Em^*(\cdot) =
\E^*(\cdot \mid \Ms)$, etc.) An important property is
that $\Ms$ is \emph{independent} of the $X_{nt}$'s, $V_{nt}$'s, and
$u_{nj}$'s, so we can treat any
$M_{nj}$ and $J_n$ terms as constants within $\Em(\cdot)$ and
$\Pm[\cdot]$ and integrate over the unconditional distributions of the
other random variables.\footnote{%
  A referee suggested the following clarification:
  \begin{itemize}
  \item[$\circ$] $\Em^*(\cdot)$ denotes the expectation over the $u_{ni}$, holding
    the sample and $M_{ni}$ fixed.
  \item[$\circ$] $\Em(\cdot)$ denotes the expectation over the sample and
    $u_{ni}$, holding only the $M_{ni}$ fixed.
  \item[$\circ$] $\E^*(\cdot)$ denotes the expectation over the $u_{ni}$ and
    $M_{ni}$, holding the sample fixed.
  \end{itemize}
  The same distinctions hold for the probabilities as well.} %
This property is especially important because it allows us to use
maximal inequalities and other moment inequalities for mixingale
arrays with only small modifications to construct almost sure bounds
on the conditional moments $\Em$.

@@

\begin{lem}\label{L2}
  If $\{A_n\}$ is a sequence of events in $\Omega$ then the following
  are equivalent:
  \begin{align*}
  \pr[A_n] &\to 0, &
  \pr^*[A_n] &\to 0 \text{\ in\ }L_1, &
  &\text{and}&
  \Pm^*[A_n] &\to 0 \text{\ in\ }L_1.
  \end{align*}
\end{lem}
\begin{proof}
  Since $\pr[A_n] = \E |\pr^*[A_n]| = \E |\pr^*_{\mathcal{M}}[A_n]|$ these
  conditions are equivalent by definition.
\end{proof}

\begin{lem}\label{L3}
Under the conditions of Theorem~\ref{T1},
\begin{equation}\label{eq:40}
  \E \Big(\sum_{j=1}^{J_n} \Big( \sum_{t = K_{n,j-1}+1}^{K_{nj}}
  (\mu_{nt}^{*} - \bar \mu_n)\Big)^2\Big) = o(n p_n)
\end{equation}
and $\bar \mu_n^{*} = \E^{*} \bar \mu_n^{*} + o_{L_2}(n^{-1/2})$.
\end{lem}
\begin{proof}
  The second result, on $\bar\mu_n^*$, is an immediate implication of
  Equation~\eqref{eq:40}. To show~\eqref{eq:40}, observe that
  \begin{align*}
    \E \Big(\sum_{j=1}^{J_n} &\Big( \sum_{t= K_{n,j-1} +1}^{K_{nj}}
    (\mu_{nt}^{*} - \bar \mu_n)\Big)^2\Big)= \\
    &= \E\Big(\sum_{j=1}^{J_n} \big\lvert \sum_{s,t = K_{n,j-1}+1}^{K_{nj}}
    \E \big((\mu_{ns}^{*} - \bar{\mu}_n)
    (\mu_{nt}^{*} - \bar{\mu}_n) \mid \Ms \big) \big\rvert \Big)\\
    &\leq \E\Big( \sum_{j=1}^{J_n}
    \sum_{k,l = -n}^{n} \tfrac{1}{n} \sum_{\tau=0}^{n-1} \big\lvert
    (\mu_{n,\tau+k} - \bar{\mu}_n) (\mu_{n,\tau+l} - \bar{\mu}_n) \;
    1\{ \tau + k, \tau + l \in I_n(\tau, M_{nj})\} \big\rvert \Big)\\
    &\leq \E\Big(\sum_{j=1}^{J_n} \Big[
    \sum_{k = -n}^{n} \big(\tfrac{1}{n} \sum_{\tau=0}^{n-1}
    (\mu_{n,\tau+k} - \bar{\mu}_n)^2
    1\{ \tau + k \in I_n(\tau, M_{nj})\}\big)^{1/2}\Big]^{2}\Big)\\
    &= \E\Big(\sum_{j=1}^{J_n}
    M_{nj}^2 \Big\lvert \tfrac{1}{n} \sum_{\tau=0}^{n-1}
    (\mu_{n,\tau} - \bar{\mu}_n)^2 \Big\lvert\Big)\\
    &= o(n p_n).
  \end{align*}
  The first equality follows from the LIE\@. The first inequality
  exploits the fact that the conditional expectation in the previous
  expression averages over the block's starting period but treats
  every other random variable as constant. (The argument of the
  indicator function and the indices of the summations are
  deliberately chosen to accommodate blocks that include observations
  from both the beginning and the end of the series.) The second
  inequality is an application of the Cauchy-Schwarz inequality to the
  inner-most summation. The next equality holds because the
  summation
  \[
    \sum_{\tau=0}^{n-1} (\mu_{n,\tau+k} - \bar{\mu}_n)^2
    1\{ \tau + k \in I_n(\tau, M_{nj})\}
  \]
  is zero for all but $M_{nj}$ values of $k$ and is a nonzero constant
  for those $k$. The last line holds by Lemma~\ref{L1} and assumption.
\end{proof}

@@

  Since $Z_{nj}^*$ is a martingale difference array, Doob's maximal
  inequality for martingales \citep[see][Theorem 15.15, for
  example]{Dav:94} along with the construction of $Z_{nj}^{*}$ gives
  a bound for the first term on the \allcaps{RHS} of~\eqref{eq:6},
  \begin{align*}
  \E \Em^* & \Big[\ind(A) \tfrac{n}{m'}\max_{m = 1,\dots, m'}
  \Big(Z_n^*(\tau, K_{n,J(\tau)} - \tau)
  + \sum_{j=J(\tau)+1}^{J(\tau + m)-1} Z_{nj}^*\Big)^2\Big] \\
  &\leq \tfrac{1}{m'} \sum_{u = 0}^{n-1} \E \Big[\Em \big[
  \ind(A) Z_n(u, K_{n,J(\tau)} - \tau)^2\big] +
  \sum_{j=J(\tau)+1}^{J(\tau + m')-1} \Em \big[\ind(A) Z_n(u, M_{nj})^2\big]\Big] \\
  &\leq \tfrac{3}{m'} \sum_{u = 0}^{n-1} \E\Big[\Em\big[
  \ind(A) Z'_n(u, K_{n,J(\tau)} - \tau)^2\big]
  + \tfrac{1}{n} \Pm[A] \Big(\sum_{t \in I_n(u, K_{n,J(\tau)} - \tau)}
  (\mu_{nt} - \bar \mu_n)\Big)^2 \\
  & \quad+ \sum_{j=J(\tau)+1}^{J(\tau + m')-1}
  \Em \big[\ind(A) Z_n'(u, M_{nj})^2]
  + \tfrac{1}{n} \Pm[A]
  \sum_{j=J(\tau)+1}^{J(\tau + m')-1}\Big( \sum_{t \in I_n(u,M_{nj})}
  (\mu_{nt} - \bar{\mu}_n) \Big)^2 \\
  & \quad+ \Em\big( \ind(A) Z_n'(u, n)^2 \big)
  \sum_{j=J(\tau)}^{J(\tau+m')-1} M_{nj}^2/n^2 \Big] \\
  &\leq (3\delta/m') \E (K_{n,J(\tau+m')-1} - \tau)
    + 3\delta \E\Big(\sum_{j=J(\tau)}^{J(\tau+m')-1} M_{nj}^2/n^2 \Big) \\
  &\quad+ \tfrac{3\delta'}{nm'} \sum_{u=0}^{n-1}\E\Big( \Big(\sum_{t \in I_n(u, K_{n,J(\tau)} - \tau)}
  (\mu_{nt} - \bar \mu_n)\Big)^2 +
  \sum_{j=J(\tau)+1}^{J(\tau + m')-1}\Big( \sum_{t \in I_n(u,M_{nj})}
  (\mu_{nt} - \bar{\mu}_n) \Big)^2 \Big) \\
  & < C_1 \delta + C_2 p_n \delta'
  \end{align*}
  for some finite constants $C_1$ and $C_2$.
  The second inequality comes from writing
  \begin{align*}
    Z_n(\tau, m) &= Z_n'(\tau, m)
    + \tfrac{1}{\sqrt{n}} \sum_{t \in I_n(\tau, m)}
    \big((\mu_{nt} - \bar \mu_n) + (\bar \mu_n - \bar X_n)\big) \\
    &= Z_n'(\tau, m)
    + \tfrac{1}{\sqrt{n}} \sum_{t \in I_n(\tau, m)} (\mu_{nt} - \bar \mu_n)
    - (m/n) Z'_n(\tau, n),
  \end{align*}
  the third follows by construction of $A$
  and the fourth follows from an argument similar to Lemma~\ref{L3}.

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
%  LocalWords:  mixingale Heyde's

>>>>

The assumption on the dispersion of the individual means,
$(\mu_{nt} - \bar\mu_n)^2$, is the same as \citepos{GoW:02} and
\citepos{GoJ:03} under their additional requirement that the
expected block length be $o(\sqrt{n})$. When
the blocks are larger than $o(\sqrt{n})$, a case that \citet{GoW:02}
and \citet{GoJ:03} do not consider, the dispersion must be smaller
than they require.
To see these implications of this assumption, it is helpful to
consider a simple example: a stochastic process with a single break of
size $\Delta_n$ at time $T_n$:
\begin{equation*}
\mu_{nt} =
\begin{cases}
  \mu & \text{if } t \leq T_n \\
  \mu + \Delta_n & \text{otherwise}.
\end{cases}
\end{equation*}
A small amount of algebra gives the relationship,
\begin{equation*}
  \sum_{t=1}^n (\mu_{nt} - \bar \mu_n)^2
  = T_n (n - T_n) \Delta_n^2 / n \equiv D_n;
\end{equation*}
we have introduced $D_n$ to minimize notation in the subsequent
discussion.  For the Moving and Circular Block Bootstraps,
Theorem~\ref{T1}'s Assumption 2 holds if $D_n = o(n / M_n^2)$, and for
the same relationship must hold for the expected block length for the
Stationary Bootstrap. If $\Delta_n = \Delta_0$ and $T_n = \pi n$ for
some constants $\Delta_0$ and $\pi$ and for all $n$, then
$D_n \approx \Delta_0^2 \, c (1 - c) \, n$ and Assumption~2 fails to
hold. These values of $\Delta_n$ and $T_n$ describe a scenario where
there is a break large enough that $\pi$ could be precisely
estimated \citep[as in][]{Bai:94}, but the break is ignored and the
bootstrap is applied to the entire dataset.

To see when Theorem~\ref{T1}'s Assumption~2 is satisfied, consider two
other parametrizations of this DGP\@. First, let $\Delta_n = \Delta_0$
and $T_n = T_0$ for some finite $T_0$. This mimics a setting where a
researcher intends to apply the bootstrap to a post-break sample, but
mistakenly includes a small number of pre-break observations as
well. In this case, Assumption 2 holds as long as
$M_n = o(\sqrt{n})$.  Similarly, suppose
$\Delta_n = \Delta / \sqrt{n}$ and $T_n = \pi n$, which mimics a
setting with a moderate break with unknown timing.  Then
$D_n \approx \Delta c (1 - c)$ and Assumption 2 again holds as long as
$M_n = o(\sqrt{n})$. Consequently, using block sizes of length
$o(\sqrt{n})$ ensures that Theorem 1's assumptions on the dispersion
of the $\mu_{nt}$ hold when there is a moderate amount of
heterogeneity---either because the sample is chosen to lie after a
large break estimated break or because any breaks are smaller in
magnitude and, consequently, potentially difficult to detect. When the
degree of heterogeneity is very small or nonexistent, so
$\sum_t (\mu_{nt} - \bar \mu_n)^2$ grows even slower, then larger
block lengths can also satisfy this condition.

@@

Our insight is that the expectation of squared blocks of the bootstrap
process can be expressed as a sequence of \emph{contiguous} blocks of
the original process, so the arguments that establish convergence of
the original squared blocks can be applied with only minor changes to
the bootstrapped blocks. Consider the Moving Blocks Bootstrap,%
\footnote{To make this presentation as simple as possible, assume for
  now that $n = M_n J_n$ exactly.} %
for example, and let
\begin{equation*}
  Z_{nj}^* = \tfrac{1}{\sqrt{n}} \sum_{t=(j-1) M_n + 1}^{j M_n} (X_{nt}^* - \bar X_n).
\end{equation*}
Then, conditional on the data, the $Z_{nj}^{*2}$ are independent can
be expected to obey an \lln, so %
$\sum_j (Z_{nj}^{*2} - \E^* Z_{nj}^{*2}) \to^p 0$ %
and the CLT for the bootstrapped array requires $\sum_j \E^*
Z_{nj}^{*2}$ to converge to a positive and finite limit.  But, since
$\E^*$ only averages over the starting point of each block, we have
\begin{align*}
  \E^* Z_{nj}^{*2} &=
  \tfrac{1}{n} \sum_{\tau = 0}^{n-M_n}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= \tau + 1}^{\tau + M_n} (X_{nt} - \bar X_n) \Big)^2 \\
  &= \tfrac{1}{n} \sum_{\tau_0 = 0}^{M_n-1}
  \sum_{j=1}^{J_n}
  \Big(\tfrac{1}{\sqrt{n}}
  \sum_{t= (j - 1) M_n + \tau_0 + 1}^{j M_n + \tau_0} (X_{nt} - \bar X_n) \Big)^2
\end{align*}
after grouping blocks separated by $M_n$ periods. For each $\tau_0$,
the summation can be expected to converge in probability through the
same arguments that were used to establish the \clt\ for the original
array.%
\footnote{Some of the details of the argument will typically need to
  change because the original \clt\ only requires convergence for
  $\tau_0 = 0$, but these details are often incidental to the original
  argument.}%
\footnote{Lemmas~\ref{L4} and~\ref{L6} are particularly
  strong demonstrations of this argument.} %
A similar representation is available for the Circular and Stationary
bootstraps.

@@

First, the relevant probability infrastructure.  Let
$(S, \Omega, \pr)$ be a probability space and define the sequence of
sub-sigma-fields $\Omega_n \subset \Omega$. Assume that each
$X_{nt}$, $V_{nt}$, $M_{nj}$, and $u_{nj}$ is $\Omega_n$-measurable,
with $u_{nj}$ the $\operatorname{uniform}(1,\dots,n)$ random variables
that designate the start period of each bootstrap ``block.'' Also
define the \sigmafield\ generated by the stationary bootstrap's block
lengths alone,
\begin{equation}
  \Ms = \sigma(J_{n}, M_{n1},\dots,M_{nJ_n})
\end{equation}
and the conditional probability $\Pm(\cdot) = \pr[\cdot \mid \Ms]$.
(And define $\Pm^*(\cdot) = \pr^*[\cdot \mid \Ms]$, $\Em(\cdot) =
\E^*(\cdot \mid \Ms)$, etc.) An important property is
that $\Ms$ is \emph{independent} of the $X_{nt}$'s, $V_{nt}$'s, and
$u_{nj}$'s, so we can treat any
$M_{nj}$ and $J_n$ terms as constants within $\Em(\cdot)$ and
$\Pm[\cdot]$ and integrate over the unconditional distributions of the
other random variables. This property is especially important because it allows us
to use maximal inequalities and other moment inequalities for
mixingale arrays with only small modifications to construct almost sure bounds on the
conditional moments $\Em$.

@@

\begin{lem}\label{L2}
  If $\{A_n\}$ is a sequence of events in $\Omega$ then the following
  are equivalent:
  \begin{align*}
  \pr[A_n] &\to 0, &
  \pr^*[A_n] &\to 0 \text{\ in\ }L_1, &
  &\text{and}&
  \Pm^*[A_n] &\to 0 \text{\ in\ }L_1.
  \end{align*}
\end{lem}
\begin{proof}
  Since $|\pr[A_n]| = \E |\pr^*[A_n]| = \E |\pr^*_{\mathcal{M}}[A_n]|$ these
  conditions are equivalent by definition.
\end{proof}

\begin{lem}\label{L3}
Under the conditions of Theorem~\ref{T1},
\begin{equation}\label{eq:40}
  \E \Big(\sum_{j=1}^{J_n} \Big( \sum_{t = K_{n,j-1}+1}^{K_{nj}}
  (\mu_{nt}^{*} - \bar \mu_n)\Big)^2\Big) = o(n p_n)
\end{equation}
and $\bar \mu_n^{*} = \E^{*} \bar \mu_n^{*} + o_{L_2}(n^{-1/2})$.
\end{lem}
\begin{proof}
  The second result, on $\bar\mu_n^*$, is an immediate implication of
  Equation~\eqref{eq:40}. To show~\eqref{eq:40}, observe that
  \begin{align*}
    \E \Big(\sum_{j=1}^{J_n} &\Big( \sum_{t= K_{n,j-1} +1}^{K_{nj}}
    (\mu_{nt}^{*} - \bar \mu_n)\Big)^2\Big)= \\
    &= \E\Big(\sum_{j=1}^{J_n} \big\lvert \sum_{s,t = K_{n,j-1}+1}^{K_{nj}}
    \E \big((\mu_{ns}^{*} - \bar{\mu}_n)
    (\mu_{nt}^{*} - \bar{\mu}_n) \mid \Ms \big) \big\rvert \Big)\\
    &\leq \E\Big( \sum_{j=1}^{J_n}
    \sum_{k,l = -n}^{n} \tfrac{1}{n} \sum_{\tau=0}^{n-1} \big\lvert
    (\mu_{n,\tau+k} - \bar{\mu}_n) (\mu_{n,\tau+l} - \bar{\mu}_n) \;
    1\{ \tau + k, \tau + l \in I_n(\tau, M_{nj})\} \big\rvert \Big)\\
    &\leq \E\Big(\sum_{j=1}^{J_n} \Big[
    \sum_{k = -n}^{n} \big(\tfrac{1}{n} \sum_{\tau=0}^{n-1}
    (\mu_{n,\tau+k} - \bar{\mu}_n)^2
    1\{ \tau + k \in I_n(\tau, M_{nj})\}\big)^{1/2}\Big]^{2}\Big)\\
    &= \E\Big(\sum_{j=1}^{J_n}
    M_{nj}^2 \Big\lvert \tfrac{1}{n} \sum_{\tau=0}^{n-1}
    (\mu_{n,\tau} - \bar{\mu}_n)^2 \Big\lvert\Big)\\
    &= o(n p_n).
  \end{align*}
  The first equality follows from the LIE\@. The first inequality
  exploits the fact that the conditional expectation in the previous
  expression averages over the block's starting period but treats
  every other random variable as contstant. (The argument of the
  indicator function and the indices of the summations are
  deliberately chosen to accomodate blocks that include observations
  from both the beginning and the end of the series.) The second
  inequality is an application of the Cauchy-Schwarz inequality to the
  inner-most summation. The next equality holds because the
  summation
  \[
    \sum_{\tau=0}^{n-1} (\mu_{n,\tau+k} - \bar{\mu}_n)^2
    1\{ \tau + k \in I_n(\tau, M_{nj})\}
  \]
  is zero for all but $M_{nj}$ values of $k$ and is a nonzero constant
  for those $k$. The last line holds by Lemma~\ref{L1} and assumption.
\end{proof}

@@

  Since $Z_{nj}^*$ is a martingale difference array, Doob's maximal
  inequality for martingales \citep[see][Theorem 15.15, for
  example]{Dav:94} along with the construction of $Z_{nj}^{*}$ gives
  a bound for the first term on the \allcaps{RHS} of~\eqref{eq:6},
  \begin{align*}
  \E \Em^* & \Big[\ind(A) \tfrac{n}{m'}\max_{m = 1,\dots, m'}
  \Big(Z_n^*(\tau, K_{n,J(\tau)} - \tau)
  + \sum_{j=J(\tau)+1}^{J(\tau + m)-1} Z_{nj}^*\Big)^2\Big] \\
  &\leq \tfrac{1}{m'} \sum_{u = 0}^{n-1} \E \Big[\Em \big[
  \ind(A) Z_n(u, K_{n,J(\tau)} - \tau)^2\big] +
  \sum_{j=J(\tau)+1}^{J(\tau + m')-1} \Em \big[\ind(A) Z_n(u, M_{nj})^2]\Big] \\
  &\leq \tfrac{3}{m'} \sum_{u = 0}^{n-1} \E\Big[\Em\big[
  \ind(A) Z'_n(u, K_{n,J(\tau)} - \tau)^2\big]
  + \tfrac{1}{n} \Pm[A] \Big(\sum_{t \in I_n(u, K_{n,J(\tau)} - \tau)}
  (\mu_{nt} - \bar \mu_n)\Big)^2 \\
  & \quad+ \sum_{j=J(\tau)+1}^{J(\tau + m')-1}
  \Em \big[\ind(A) Z_n'(u, M_{nj})^2]
  + \tfrac{1}{n} \Pm[A]
  \sum_{j=J(\tau)+1}^{J(\tau + m')-1}\Big( \sum_{t \in I_n(u,M_{nj})}
  (\mu_{nt} - \bar{\mu}_n) \Big)^2 \\
  & \quad+ \Em\big( \ind(A) Z_n'(u, n)^2 \big)
  \sum_{j=J(\tau)}^{J(\tau+m')-1} M_{nj}^2/n^2 \Big] \\
  &\leq (3\delta/m') \E (K_{n,J(\tau+m')-1} - \tau)
    + 3\delta \E\Big(\sum_{j=J(\tau)}^{J(\tau+m')-1} M_{nj}^2/n^2 \Big) \\
  &\quad+ \tfrac{3\delta'}{nm'} \sum_{u=0}^{n-1}\E\Big( \Big(\sum_{t \in I_n(u, K_{n,J(\tau)} - \tau)}
  (\mu_{nt} - \bar \mu_n)\Big)^2 +
  \sum_{j=J(\tau)+1}^{J(\tau + m')-1}\Big( \sum_{t \in I_n(u,M_{nj})}
  (\mu_{nt} - \bar{\mu}_n) \Big)^2 \Big) \\
  & < C_1 \delta + C_2 p_n \delta'
  \end{align*}
  for some finite constants $C_1$ and $C_2$.
  The second inequality comes from writing
  \begin{align*}
    Z_n(\tau, m) &= Z_n'(\tau, m)
    + \tfrac{1}{\sqrt{n}} \sum_{t \in I_n(\tau, m)}
    \big((\mu_{nt} - \bar \mu_n) + (\bar \mu_n - \bar X_n)\big) \\
    &= Z_n'(\tau, m)
    + \tfrac{1}{\sqrt{n}} \sum_{t \in I_n(\tau, m)} (\mu_{nt} - \bar \mu_n)
    - (m/n) Z'_n(\tau, n),
  \end{align*}
  the third follows by construction of $A$
  and the fourth follows from an argument similar to Lemma~\ref{L3}.

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
